### 2021-10-27 paper.tex
initial commit

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Journal Article
% LaTeX Template
% Version 1.4 (15/5/16)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Frits Wenneker (http://www.howtotex.com) with extensive modifications by
% Vel (vel@LaTeXTemplates.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[twoside,twocolumn]{article}

\usepackage{blindtext} % Package to generate dummy text throughout this template

\usepackage[sc]{mathpazo} % Use the Palatino font
\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
\linespread{1.05} % Line spacing - Palatino needs more space between lines
\usepackage{microtype} % Slightly tweak font spacing for aesthetics

\usepackage[english]{babel} % Language hyphenation and typographical rules
\usepackage{graphicx}

\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry} % Document margins
\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
\usepackage{booktabs} % Horizontal rules in tables

\usepackage{lettrine} % The lettrine is the first enlarged letter at the beginning of the text

\usepackage{enumitem} % Customized lists
\setlist[itemize]{noitemsep} % Make itemize lists more compact

\usepackage{abstract} % Allows abstract customization
\renewcommand{\abstractnamefont}{\normalfont\bfseries} % Set the "Abstract" text to bold
\renewcommand{\abstracttextfont}{\normalfont\small\itshape} % Set the abstract itself to small italic text

\usepackage{titlesec} % Allows customization of titles
\renewcommand\thesection{\Roman{section}} % Roman numerals for the sections
\renewcommand\thesubsection{\roman{subsection}} % roman numerals for subsections
\titleformat{\section}[block]{\large\scshape\centering}{\thesection.}{1em}{} % Change the look of the section titles
\titleformat{\subsection}[block]{\large}{\thesubsection.}{1em}{} % Change the look of the section titles

\usepackage{fancyhdr} % Headers and footers
\pagestyle{fancy} % All pages have headers and footers
\fancyhead{} % Blank out the default header
\fancyfoot{} % Blank out the default footer
\fancyhead[C]{Running title $\bullet$ May 2016 $\bullet$ Vol. XXI, No. 1} % Custom header text
\fancyfoot[RO,LE]{\thepage} % Custom footer text

\usepackage{titling} % Customizing the title section

\usepackage{hyperref} % For hyperlinks in the PDF

%----------------------------------------------------------------------------------------
%	TITLE SECTION
%----------------------------------------------------------------------------------------

\setlength{\droptitle}{-4\baselineskip} % Move the title up

\pretitle{\begin{center}\Huge\bfseries} % Article title formatting
\posttitle{\end{center}} % Article title closing formatting
\title{Baysian Optimization in MPC} % Article title
\author{%
\textsc{Frederic Dlugi}\\[1ex] % Your name
\normalsize Universität zu Lübeck \\ % Your institution
\normalsize \href{mailto:frederic.dlugi@student.uni-luebeck.de}{frederic.dlugi@student.uni-luebeck.de} % Your email address
%\and % Uncomment if 2 authors are required, duplicate these 4 lines if more
%\textsc{Jane Smith}\thanks{Corresponding author} \\[1ex] % Second author's name
%\normalsize University of Utah \\ % Second author's institution
%\normalsize \href{mailto:jane@smith.com}{jane@smith.com} % Second author's email address
}
\date{\today} % Leave empty to omit a date
\renewcommand{\maketitlehookd}{%
\begin{abstract}
\noindent \blindtext % Dummy abstract text - replace \blindtext with your abstract text
\end{abstract}
}

%----------------------------------------------------------------------------------------

\begin{document}

% Print the title
\maketitle

%----------------------------------------------------------------------------------------
%	ARTICLE CONTENTS
%----------------------------------------------------------------------------------------

\section{Introduction}

\lettrine[nindent=0em,lines=3]{W} e start by introducing bayesian optimization and MPC on a high level.
This is done to let you know what I understand when talking about these algorithms.

\subsection{What is Bayesian Optimization?}
Bayesian optimization is a strategy for global optimization of backbox functions.
It therefore does not require the computation of gradients, but works best on continuous functions.
It is best suited for optimizing functions, where each evaluation takes a long time.
The number of input dimensions for bayesian optimization is typically less then $20$.\cite{frazier2018tutorial}

\subsection{What is MPC?}
MPC is an acronym for model predicive control.
It replaces classical control algorithms that usually work in an offline manner with an optimizer,
that solves the optimal control problem for a receding horizon (Figure \ref{fig:mpc}).
Each action taken is the first action of the optimal control plan.
The plan gets updated continuously (Figure \ref{fig:mpc_plan}).

\begin{figure}[h]
    \caption{Blockdiagram of MPC algorithm}
    \centering
    \includegraphics[width=0.45\textwidth]{fig_mpc.pdf}
    \label{fig:mpc}
\end{figure}

\begin{figure}[h]
    \caption{Plan horizon of MPC algorithm}
    \centering
    \includegraphics[width=0.45\textwidth]{fig_mpc_plan.pdf}
    \label{fig:mpc_plan}
\end{figure}

%------------------------------------------------

\section{Methods}

Maecenas sed ultricies felis. Sed imperdiet dictum arcu a egestas.
\begin{itemize}
\item Donec dolor arcu, rutrum id molestie in, viverra sed diam
\item Curabitur feugiat
\item turpis sed auctor facilisis
\item arcu eros accumsan lorem, at posuere mi diam sit amet tortor
\item Fusce fermentum, mi sit amet euismod rutrum
\item sem lorem molestie diam, iaculis aliquet sapien tortor non nisi
\item Pellentesque bibendum pretium aliquet
\end{itemize}
\blindtext % Dummy text

Text requiring further explanation\footnote{Example footnote}.

%------------------------------------------------

\section{Results}

\begin{table}
\caption{Example table}
\centering
\begin{tabular}{llr}
\toprule
\multicolumn{2}{c}{Name} \\
\cmidrule(r){1-2}
First name & Last Name & Grade \\
\midrule
John & Doe & $7.5$ \\
Richard & Miles & $2$ \\
\bottomrule
\end{tabular}
\end{table}

\blindtext % Dummy text

\begin{equation}
\label{eq:emc}
e = mc^2
\end{equation}

\blindtext % Dummy text

%------------------------------------------------

\section{Discussion}

\subsection{Subsection One}

A statement requiring citation \cite{sorourifar2021data}.
\blindtext % Dummy text

\subsection{Subsection Two}

\blindtext % Dummy text

%----------------------------------------------------------------------------------------
%	REFERENCE LIST
%----------------------------------------------------------------------------------------
\bibliographystyle{plain}
\bibliography{refs.bib}
%----------------------------------------------------------------------------------------

\end{document}

### 2021-11-26 src/controller.py
add code

@@ -1,0 +1,47 @@
+
+from models.kinematric_bicycle_model import Bicycle
+import numpy as np
+import copy
+from scipy.optimize import minimize
+
+class MpcController:
+    def __init__(self):
+        self.horiz = None
+        self.R = np.diag([0.01, 0.01])                 # input cost matrix
+        # input difference cost matrix
+        self.Rd = np.diag([0.01, 0.01])
+        self.Q = np.diag([1.0, 1.0])                   # state cost matrix
+        self.Qf = self.Q                               # state final matrix
+
+    def cost(
+            self,
+            u_k: np.ndarray,
+            my_car: Bicycle,
+            points: np.ndarray):
+        mpc_car = copy.copy(my_car)
+        u_k = u_k.reshape((self.horiz, 2)).T
+        z_k = np.zeros((2, self.horiz + 1))
+
+        desired_state = points.T
+        cost = 0.0
+
+        for i in range(self.horiz):
+            mpc_car.step(u_k[0, i], u_k[1, i])
+
+            z_k[0, i] = mpc_car.xc
+            z_k[1, i] = mpc_car.yc
+            cost += np.sum(self.R @ (u_k[:, i]**2))
+            cost += np.sum(self.Q @ ((desired_state[:, i] - z_k[:, i])**2))
+            if i < (self.horiz - 1):
+                cost += np.sum(self.Rd @ ((u_k[:, i + 1] - u_k[:, i])**2))
+        return cost
+
+    def optimize(self, my_car: Bicycle, points: np.ndarray):
+        self.horiz = points.shape[0]
+        bnd = [(-5, 5), (-1.22, 1.22)] * self.horiz
+        result = minimize(
+            self.cost, args=(
+                my_car, points), x0=np.zeros(
+                (2 * self.horiz)), method='SLSQP', bounds=bnd)
+        #print(result)
+        return result.x[0], result.x[1]

### 2021-11-26 src/main.py

@@ -1,0 +1,18 @@
+from simulation import run_sim
+from bayes_opt import BayesianOptimization
+from matplotlib import pyplot as plt
+import numpy as np
+
+def plot_bo(f, bo):
+    x = np.linspace(-2, 10, 10000)
+    mean, sigma = bo._gp.predict(x.reshape(-1, 1), return_std=True)
+
+    plt.figure(figsize=(16, 9))
+    plt.plot(x, f(x))
+    plt.plot(x, mean)
+    plt.fill_between(x, mean + sigma, mean - sigma, alpha=0.1)
+    plt.scatter(bo.space.params.flatten(), bo.space.target, c="red", s=50, zorder=10)
+    plt.show()
+
+bo = BayesianOptimization(f=run_sim, pbounds={"lr":(0.5,1.5),"lf":(0.5,1.5)}, random_state=1)
+print(bo.maximize(n_iter=50))

### 2021-11-26 src/models/kinematric_bicycle_model.py

@@ -1,0 +1,40 @@
+# https://github.com/daniel-s-ingram/self_driving_cars_specialization/blob/master/1_introduction_to_self_driving_cars/Kinematic_Bicycle_Model.ipynb
+
+from math import sin, cos, tan, atan2
+import numpy as np
+
+class Bicycle():
+    def __init__(self, lr=1.2, lf=0.8, dt=0.01):
+        self.xc = 0
+        self.yc = 0
+        self.theta = 0
+        self.delta = 0
+        self.beta = 0
+
+        self.L = lf + lr
+        self.lr = lr
+        self.w_max = 1.22
+
+        self.dt = dt
+
+    def reset(self):
+        self.xc = 0
+        self.yc = 0
+        self.theta = 0
+        self.delta = 0
+        self.beta = 0
+
+    def step(self, v, w):
+        xc_dot = v * cos(self.theta + self.beta)
+        yc_dot = v * sin(self.theta + self.beta)
+        theta_dot = v * cos(self.beta) * tan(self.delta) / self.L
+        delta_dot = max(-self.w_max, min(self.w_max, w))
+        self.xc += xc_dot * self.dt
+        self.yc += yc_dot * self.dt
+        self.theta += theta_dot * self.dt
+        self.delta += delta_dot * self.dt
+        self.beta = atan2(self.lr * tan(self.delta), self.L)
+
+    @property
+    def state(self):
+        return np.array([self.xc, self.yc, self.theta, self.delta, self.beta])

### 2021-11-26 src/path/path_generation.py

@@ -1,0 +1,20 @@
+from typing import Tuple
+import numpy as np
+from scipy import interpolate
+from numpy import random
+
+def generate_random_path(num_waypoints: int, bounds: Tuple[float], num_steps):
+    waypoints = random.random((num_waypoints, 2))
+    waypoints *= (bounds[1] - bounds[0])
+    waypoints += bounds[0]
+    waypoints[-1,:] = waypoints[0,:]
+
+    return generate_path(waypoints, num_steps)
+
+
+def generate_path(waypoints: np.ndarray, num_steps: int):
+    tck, _ = interpolate.splprep([waypoints[:, 0], waypoints[:, 1]], s=0)
+    x = np.linspace(0, 1, num=num_steps, endpoint=True)
+    out = interpolate.splev(x, tck)
+
+    return np.array(out).T

### 2021-11-26 src/simulation.py

@@ -1,0 +1,136 @@
+
+from models.kinematric_bicycle_model import Bicycle
+from controller import MpcController
+from path.path_generation import generate_random_path
+from matplotlib import pyplot as plt
+import numpy as np
+from numpy import random
+import copy
+import pathlib
+import warnings
+warnings.filterwarnings("ignore", category=RuntimeWarning)
+
+ref = [[25082, 20.532948320013894],
+        [53816, 210.4562569750449],
+        [76064, 55.96925996037113],
+        [58549, 14.468911686067464],
+        [427, 184.98888084734335],
+        [16056, 60.090698631201114],
+        [99010, 301.4617037514949],
+        [88390, 6.390730425176852],
+        [75542, 32.69409353462841],
+        [30399, 81.33519708084226],
+        [7108, 19.139958767371365],
+        [3726, 14.942495462629457],
+        [95374, 9.733050214245466],
+        [50720, 38.02823104644416],
+        [77323, 9.824426545346814],
+        [6439, 74.75510398852445],
+        [4875, 13.24150741589534],
+        [3608, 400.3107905577367],
+        [12763, 49.95084762012596],
+        [41622, 98.4370625781701],
+        [58798, 4.801656390614494],
+        [35702, 312.3905652160304],
+        [18806, 4.288301349816508],
+        [36082, 49.36674991612697],
+        [89159, 41.15665530305945],
+        [2686, 31.61139954721623],
+        [54004, 28.732657935459557],
+        [50386, 333.87523188216505],
+        [85580, 131.3118458004769],
+        [77845, 370.71044711132976],
+        [56685, 302.9659089976724],
+        [5792, 5.755681452443642],
+        [79745, 50.04276967431635],
+        [93168, 32.22235967555776],
+        [13279, 53.334707673612115],
+        [89706, 618.0473584210496],
+        [704, 18.677325007211916],
+        [18736, 217.16298237934285],
+        [87677, 15.615496317709164],
+        [51454, 25.292586051428195],
+        [96712, 25.823740309285],
+        [90297, 4.843383774860714],
+        [38163, 67.23338098787063],
+        [41889, 59.88144258006616],
+        [79491, 46.566620641858094],
[…]

### 2021-11-26 src/models/kinematric_bicycle_model.py
overwrite mpc_bike state with real state

@@ -38,3 +38,7 @@ class Bicycle():
     @property
     def state(self):
         return np.array([self.xc, self.yc, self.theta, self.delta, self.beta])
+
+    @state.setter
+    def state(self, value):
+        self.xc, self.yc, self.theta, self.delta, self.beta = value

### 2021-11-26 src/simulation.py

@@ -92,8 +92,8 @@ def run_sim(lr=1.2, lf=0.8):
 
         for i, _ in enumerate(path):
             v, w = controller.optimize(mpc_bike, path[i:i + MPC_HORIZON,:])
-            mpc_bike.step(v, w)
             bike.step(v, w)
+            mpc_bike.state = bike.state
 
             u[i,:] += [v, w]
 

### 2021-11-26 src/simulation.py
overwrite only position and not whole state

@@ -93,12 +93,14 @@ def run_sim(lr=1.2, lf=0.8):
         for i, _ in enumerate(path):
             v, w = controller.optimize(mpc_bike, path[i:i + MPC_HORIZON,:])
             bike.step(v, w)
-            mpc_bike.state = bike.state
+            mpc_bike.step(v, w)
 
             u[i,:] += [v, w]
 
             pos[i,:] = [bike.xc, bike.yc]
             pos_c[i,:] = [mpc_bike.xc, mpc_bike.yc]
+            mpc_bike.xc = bike.xc
+            mpc_bike.yc = bike.yc
 
             # show control output
             plt.cla()

### 2021-11-27 src/simulation.py
change training tracks

@@ -62,7 +62,7 @@ ref = [[25082, 20.532948320013894],
         [41673, 8.70336298911908]]
 
 def run_sim(lr=1.2, lf=0.8):
-    train_tracks = [58549]#, 25082, 53816, 76064, 427, 16056, 99010, 88390, 75542, 30399, 7108, 3726, 95374, 50720, 77323, 6439, 4875, 3608, 12763, 41622, 58798, 35702, 18806, 36082, 89159, 2686, 54004, 50386, 85580, 77845, 56685, 5792, 79745, 93168, 13279, 89706, 704, 18736, 87677, 51454, 96712, 90297, 38163, 41889, 79491, 36556, 19014, 25424, 92508, 41673]
+    train_tracks = [58798, 18806, 90297]#[58549, 25082, 53816, 76064, 427, 16056, 99010, 88390, 75542, 30399, 7108, 3726, 95374, 50720, 77323, 6439, 4875, 3608, 12763, 41622, 58798, 35702, 18806, 36082, 89159, 2686, 54004, 50386, 85580, 77845, 56685, 5792, 79745, 93168, 13279, 89706, 704, 18736, 87677, 51454, 96712, 90297, 38163, 41889, 79491, 36556, 19014, 25424, 92508, 41673]
     MPC_HORIZON = 10
     plt.ion()
     plt.show()

### 2021-12-16 src/main.py
update log files add log visualization update cli for main script

@@ -24,18 +24,21 @@ def plot_bo(f, bo):
 
 parser = argparse.ArgumentParser(
     description="Run BO on an MPC problem to find L_f and L_r params.")
+
 parser.add_argument(
     "--target_lf",
     "-tlf",
     help="Target value for L_f",
     default=1.2,
     type=float)
+
 parser.add_argument(
     "--target_lr",
     "-tlr",
     help="Target value for L_r",
     default=0.8,
     type=float)
+
 parser.add_argument(
     "--bounds",
     "-b",
@@ -45,11 +48,13 @@ parser.add_argument(
         0.5,
         1.5],
     type=float)
+
 parser.add_argument(
     "--n_iter",
     help="Number of iterations of BO",
     default=50,
     type=int)
+
 parser.add_argument(
     "--acq",
     help="Acquisition function of BO (ei, ucb or poi)",
@@ -70,14 +75,48 @@ parser.add_argument(
     "--measured_states",
     help="Measure theses states from actual controller (0-4)",
     nargs="*",
-    default=[0,1],
+    default=[0, 1],
     type=int)
 
 parser.add_argument(
-    "--num_references",
-    help="Number of references per evaluation",
+    "--track_waypoints",
+    help="Number of waypoints per track",
+    default=7,
+    type=int)
+
+parser.add_argument(
+    "--track_bounds",
+    nargs=2,
+    help="Track bounds",
+    default=[
+        0.,
+        5.],
+    type=float)
+
+parser.add_argument(
+    "--track_num_points",
+    help="Number of points per track",
+    default=200,
+    type=int)
+
+parser.add_argument(
+    "--track_seeds",
+    help="Seeds of points per track",
+    nargs="+",
+    default=[18806, 58798, 90297],
+    type=int)
+
+parser.add_argument(
+    "--track_num",
+    help="Number of tracks per evaluation (ignored if track_seeds != default)",
     default=3,
     type=int)
+
+parser.add_argument(
+    "--gp_alpha",
+    help="Alpha value for GP",
+    default=1e-6,
+    type=float)
 
 args = parser.parse_args()
 
@@ -85,10 +124,16 @@ bounds = (args.bounds[0], args.bounds[1])
 
 measured_states = args.measured_states
 
[…]

### 2021-12-16 src/path/path_generation.py

@@ -4,17 +4,17 @@ from scipy import interpolate
 from numpy import random
 
 
-def generate_random_path(num_waypoints: int, bounds: Tuple[float], num_steps):
+def generate_random_path(num_waypoints: int, bounds: Tuple[float], num_steps, s=0.):
     waypoints = random.random((num_waypoints, 2))
     waypoints *= (bounds[1] - bounds[0])
     waypoints += bounds[0]
     waypoints[-1, :] = waypoints[0, :]
 
-    return generate_path(waypoints, num_steps)
+    return generate_path(waypoints, num_steps, s)
 
 
-def generate_path(waypoints: np.ndarray, num_steps: int):
-    tck, _ = interpolate.splprep([waypoints[:, 0], waypoints[:, 1]], s=0)
+def generate_path(waypoints: np.ndarray, num_steps: int, s=0.):
+    tck, _ = interpolate.splprep([waypoints[:, 0], waypoints[:, 1]], s=s)
     x = np.linspace(0, 1, num=num_steps, endpoint=True)
     out = interpolate.splev(x, tck)
 

### 2021-12-16 src/simulation.py

@@ -11,75 +11,86 @@ import warnings
 warnings.filterwarnings("ignore", category=RuntimeWarning)
 
 reference = np.array([[25082, 20.532948320013894],
-             [53816, 210.4562569750449],
-             [76064, 55.96925996037113],
-             [58549, 14.468911686067464],
-             [427, 184.98888084734335],
-             [16056, 60.090698631201114],
-             [99010, 301.4617037514949],
-             [88390, 6.390730425176852],
-             [75542, 32.69409353462841],
-             [30399, 81.33519708084226],
-             [7108, 19.139958767371365],
-             [3726, 14.942495462629457],
-             [95374, 9.733050214245466],
-             [50720, 38.02823104644416],
-             [77323, 9.824426545346814],
-             [6439, 74.75510398852445],
-             [4875, 13.24150741589534],
-             [3608, 400.3107905577367],
-             [12763, 49.95084762012596],
-             [41622, 98.4370625781701],
-             [58798, 4.801656390614494],
-             [35702, 312.3905652160304],
-             [18806, 4.288301349816508],
-             [36082, 49.36674991612697],
-             [89159, 41.15665530305945],
-             [2686, 31.61139954721623],
-             [54004, 28.732657935459557],
-             [50386, 333.87523188216505],
-             [85580, 131.3118458004769],
-             [77845, 370.71044711132976],
-             [56685, 302.9659089976724],
-             [5792, 5.755681452443642],
-             [79745, 50.04276967431635],
-             [93168, 32.22235967555776],
-             [13279, 53.334707673612115],
-             [89706, 618.0473584210496],
-             [704, 18.677325007211916],
-             [18736, 217.16298237934285],
-             [87677, 15.615496317709164],
-             [51454, 25.292586051428195],
-             [96712, 25.823740309285],
-             [90297, 4.843383774860714],
-             [38163, 67.23338098787063],
-             [41889, 59.88144258006616],
-             [79491, 46.566620641858094],
-             [36556, 88.76545573558323],
-             [19014, 49.39164390714606],
-             [25424, 12.838836167874437],
-             [92508, 61.953038248897045],
-             [41673, 8.70336298911908]])
+                      [53816, 210.4562569750449],
+                      [76064, 55.96925996037113],
+                      [58549, 14.468911686067464],
+                      [427, 184.98888084734335],
+                      [16056, 60.090698631201114],
+                      [99010, 301.4617037514949],
[…]

### 2021-12-16 src/visualize_log.py

@@ -1,0 +1,74 @@
+import numpy as np
+import argparse
+from os import path
+from itertools import product
+
+from sklearn.gaussian_process.kernels import Matern
+from sklearn.gaussian_process import GaussianProcessRegressor
+
+from matplotlib import pyplot as plt
+
+if __name__ == "__main__":
+    bounds = (0.5, 1.5)
+    target = [1.2, 0.8]
+    acquisition_function = "ei"
+    xi = 0
+    kappa = 2.576
+    measured_states = [0, 1]
+    num_references = 3
+    alpha = 1e-6
+
+    parser = argparse.ArgumentParser(prog="Log visualizer")
+    parser.add_argument("path_in", type=str)
+
+    args = parser.parse_args()
+
+    log_path = path.join(args.path_in, "log.txt")
+
+    log_lines = []
+    with open(log_path, 'r') as log_file:
+        log_lines = log_file.readlines()
+
+    command = log_lines[0]
+    log_lines = log_lines[3:]
+
+    evaluations = []
+    for line in log_lines:
+        if line.startswith("|"):
+            line_cells = line.split("|")
+            iteration = int(line_cells[1])
+            target = float(line_cells[2].replace("+0 ","3"))
+            lf = float(line_cells[3])
+            lr = float(line_cells[4])
+            evaluations += [[iteration, target, lf, lr]]
+
+    evaluations = np.array(evaluations)
+    best_evaluation = np.argmax(evaluations[:, 1])
+
+    gp = GaussianProcessRegressor(
+            kernel=Matern(nu=2.5),
+            alpha=1,
+            normalize_y=True,
+            n_restarts_optimizer=15,
+            random_state=42
+        )
+
+    lr_space = np.linspace(bounds[0], bounds[1])
+    lf_space = np.linspace(bounds[0], bounds[1])
+    x1x2 = np.array(list(product(lr_space, lf_space)))
+
+    gp.fit(evaluations[:, [2,3]], evaluations[:, 1])
+    y_pred = gp.predict(x1x2)
+
+    Zp = np.reshape(y_pred,(50,50))
+
+    X0p, X1p = x1x2[:,0].reshape(50,50), x1x2[:,1].reshape(50,50)
+    plt.title(command)
+    plt.contourf(X0p, X1p, Zp, levels=32)
+    plt.xlabel("$L_r$")
+    plt.ylabel("$L_f$")
[…]

### 2021-12-16 paper.tex
move papers and update gitignore
update paper

@@
 \usepackage{lettrine} % The lettrine is the first enlarged letter at the beginning of the text
+\usepackage{algorithm}
+\usepackage[noend]{algpseudocode}
+\usepackage{amsmath}
 
@@
 \begin{abstract}
-\noindent \blindtext % Dummy abstract text - replace \blindtext with your abstract text
+\noindent When using MPC to control a vehicle it is often neccesary to fine tune vehicle parameter for the model to match reality.
+This is a time consuming process that often relies on trial and error or grid search of the parameter space.
+In this paper we evaluate the use of Bayesian Optimization to tackle this problem.
+We simulate vehicle dynamics of a simple bicycle model with two parameters.
+We try to find the real vehicle parameters by optimizing controller performance with different parameters using Bayesian Optimization.
 \end{abstract}
@@
 \subsection{What is Bayesian Optimization?}
-Bayesian optimization is a strategy for global optimization of backbox functions.
+Bayesian optimization is a strategy for global optimization of backbox functions. (Alg. \ref{alg:bo})
 It therefore does not require the computation of gradients, but works best on continuous functions.
 It is best suited for optimizing functions, where each evaluation takes a long time.
-The number of input dimensions for bayesian optimization is typically less then $20$.\cite{frazier2018tutorial}
+The number of input dimensions for bayesian optimization is typically less then $20$. \cite{frazier2018tutorial}
+Bayesian Optimization uses an aquisition function that operates on a gaussian process of the evaluations.
+In this paper we evaluate three different acquisition functions: Expected Improvement (Alg.
+\ref{alg:ei}), Probability of Improvement (Alg.
+\ref{alg:poi}) and Upper Confidence Bound (Alg.
+\ref{alg:ucb}).
+We use the Matern ($\nu=2.5$) kernel for all experiments.
+We also vary the $\alpha$ Parameter of the Gaussian Process to smooth out the cost landscape.
+
+\begin{algorithm}
+    \caption{Bayesian Optimization}
+    \label{alg:bo}
+    \begin{algorithmic}
+        \State $\text{f} \gets \text{black box function}$
+        \State $\text{initPos} \gets \text{initial positions}$
+        \State $\text{evaluations} \gets \text{f(initPos)}$
+        \For{$i= 1 \to N$}
+            \State $\alpha \gets \text{acquisitionFunction(evaluations)}$
+            \State $\text{nextPos} \gets \text{argmax(}\alpha\text{)}$
+            \State $\text{evaluations}.append(\text{f(nextPos)})$
+        \EndFor
+        \State $result \gets \text{max(evaluations)}$
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}
+    \caption{Expected Improvement}
+    \label{alg:ei}
+    \begin{algorithmic}
+        \State $\mu, \sigma \gets \text{gp.predict(x)}$
+        \State $z \gets (\mu - y_{max} - xi)/\sigma$
+        \State $result \gets (\mu - y_{max} - xi) * \text{cdf(z)} + \sigma * \text{pdf(z)}$
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}
+    \caption{Probability of Improvement}
+    \label{alg:poi}
+    \begin{algorithmic}
+        \State $\mu, \sigma \gets \text{gp.predict(x)}$
+        \State $z \gets (\mu - y_{max} - xi)/\sigma$
+        \State $result \gets \text{cdf(z)}$
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}
+    \caption{Upper Confidence Bound}
+    \label{alg:ucb}
+    \begin{algorithmic}
+        \State $\mu, \sigma \gets \text{gp.predict(x)}$
+        \State $result \gets \mu + \kappa * \sigma$
+    \end{algorithmic}
+\end{algorithm}
 
@@
 MPC is an acronym for model predicive control.
-It replaces classical control algorithms that usually work in an offline manner with an optimizer,
-that solves the optimal control problem for a receding horizon (Figure \ref{fig:mpc}).
+It replaces classical control algorithms that usually work in an offline manner with an optimizer, that solves the optimal control problem for a receding horizon (Figure \ref{fig:mpc}).
 Each action taken is the first action of the optimal control plan.
-The plan gets updated continuously (Figure \ref{fig:mpc_plan}).
+The plan gets updated continuously (Figure \ref{fig:mpc_plan}), this is called receding horizon approach.
 
@@
 
-%------------------------------------------------
+\subsection{Kinematic Bicycle Model}
 
-\section{Methods}
+We use a kinematic bicycle model defined by algorithm \ref{alg:kbm} and \ref{alg:ubm}. We try to estimate the $L_r$ and $L_f$ in this paper.
+\begin{algorithm}
+    \caption{Kinematic Bicycle Model}
+    \label{alg:kbm}
+    \begin{algorithmic}
+        \State $v \gets \text{speed (input)}$
+        \State $\omega \gets \text{steering angle rate (input)}$
+        \State $x_c, y_c \gets \text{center position}$
+        \State $\theta \gets \text{absolute angle}$
+        \State $\delta \gets \text{steering angle}$
+        \State $\beta \gets \text{side slip factor}$
+        \State $L_r \gets \text{Distance center to rear wheel}$
+        \State $L_f \gets \text{Distance center to font wheel}$
+        \State $L \gets L_r + L_f$
+    \end{algorithmic}
+\end{algorithm}
 
-Maecenas sed ultricies felis. Sed imperdiet dictum arcu a egestas.
-\begin{itemize}
-\item Donec dolor arcu, rutrum id molestie in, viverra sed diam
-\item Curabitur feugiat
-\item turpis sed auctor facilisis
-\item arcu eros accumsan lorem, at posuere mi diam sit amet tortor
-\item Fusce fermentum, mi sit amet euismod rutrum
-\item sem lorem molestie diam, iaculis aliquet sapien tortor non nisi
-\item Pellentesque bibendum pretium aliquet
-\end{itemize}
-\blindtext % Dummy text
+\begin{algorithm}
+    \caption{Update Bicycle Model}
+    \label{alg:ubm}
+    \begin{algorithmic}
+        \State $\dot{x_c} \gets v * \cos{(\theta + \beta)}$
+        \State $\dot{y_c} \gets v * \sin{(\theta + \beta)}$
+        \State $\dot{\theta} \gets v * \cos{(\beta)} * \tan{(\delta)} / L$
+        \State $\dot{\delta} \gets \omega$
+        \State $\beta \gets \text{atan2}(L_r * \tan{(\delta)}, L)$
+    \end{algorithmic}
+\end{algorithm}
 
-Text requiring further explanation\footnote{Example footnote}.
+\section{Experiments}
 
-%------------------------------------------------
+Experiment Setup
 
-\section{Results}
-
-\begin{table}
-\caption{Example table}
-\centering
-\begin{tabular}{llr}
-\toprule
-\multicolumn{2}{c}{Name} \\
-\cmidrule(r){1-2}
-First name & Last Name & Grade \\
-\midrule
-John & Doe & $7.5$ \\
-Richard & Miles & $2$ \\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\blindtext % Dummy text
-
-\begin{equation}
-\label{eq:emc}
-e = mc^2
-\end{equation}
-
-\blindtext % Dummy text
-
-%------------------------------------------------
-
-\section{Discussion}
-
-\subsection{Subsection One}
-
-A statement requiring citation \cite{sorourifar2021data}.
-\blindtext % Dummy text
-
-\subsection{Subsection Two}
-
-\blindtext % Dummy text
-
-%----------------------------------------------------------------------------------------
-%	REFERENCE LIST
-%----------------------------------------------------------------------------------------
 \bibliographystyle{plain}
 \bibliography{refs.bib}
-%----------------------------------------------------------------------------------------
 

### 2022-01-23 build/lib/bo_mpc/path/path_generation.py
fix stuff and restructure

@@ -1,0 +1,30 @@
+from typing import Tuple
+import numpy as np
+from scipy import interpolate
+from numpy import random
+
+
+def generate_random_path(num_waypoints: int, bounds: Tuple[float], num_steps, s=0.):
+    waypoints = random.random((num_waypoints, 2))
+    waypoints *= (bounds[1] - bounds[0])
+    waypoints += bounds[0]
+    waypoints[-1, :] = waypoints[0, :]
+
+    return generate_path(waypoints, num_steps, s)
+
+
+def generate_path(waypoints: np.ndarray, num_steps: int, s=0.):
+    tck, _ = interpolate.splprep([waypoints[:, 0], waypoints[:, 1]], s=s)
+    x = np.linspace(0, 1, num=num_steps, endpoint=True)
+    out = interpolate.splev(x, tck)
+
+    return np.array(out).T
+
+def generate_path(bike, num_steps: int):
+    inputs = random.normal(size=(num_steps - 1, 2))
+    path = np.zeros((num_steps, 2))
+    bike.reset()
+    for i in range(1, num_steps):
+        bike.step(* inputs[i - 1, :])
+        path[i, :] = [bike.xc, bike.yc]
+    return path

### 2022-01-23 src/bo_mpc/models/kinematric_bicycle_model.py

@@ -1,0 +1,51 @@
+# https://github.com/daniel-s-ingram/self_driving_cars_specialization/blob/master/1_introduction_to_self_driving_cars/Kinematic_Bicycle_Model.ipynb
+
+from math import sin, cos, tan, atan2
+import numpy as np
+
+
+class Bicycle:
+    def __init__(self, lr=1.2, lf=0.8, dt=0.01):
+        self.reset()
+
+        self.L = lf + lr
+        self.lr = lr
+        self.a_max = 5
+        self.w_max = 1.22
+
+        self.dt = dt
+
+    def reset(self):
+        self.v = 0
+        self.xc = 0
+        self.yc = 0
+        self.theta = 0
+        self.delta = 0
+        self.beta = 0
+
+    def step(self, a, w):
+        """
+        param a:
+        accelaration
+        param w:
+        steering change
+        """
+        v_dot = max(-self.a_max, min(a, self.a_max))
+        xc_dot = self.v * cos(self.theta + self.beta)
+        yc_dot = self.v * sin(self.theta + self.beta)
+        theta_dot = self.v * cos(self.beta) * tan(self.delta) / self.L
+        delta_dot = max(-self.w_max, min(self.w_max, w))
+        self.v += v_dot * self.dt
+        self.xc += xc_dot * self.dt
+        self.yc += yc_dot * self.dt
+        self.theta += theta_dot * self.dt
+        self.delta += delta_dot * self.dt
+        self.beta = atan2(self.lr * tan(self.delta), self.L)
+
+    @property
+    def state(self):
+        return np.array([self.v, self.xc, self.yc, self.theta, self.delta, self.beta]).astype(float)
+
+    @state.setter
+    def state(self, value):
+        self.v, self.xc, self.yc, self.theta, self.delta, self.beta = value

### 2022-01-23 src/bo_mpc/simulation.py

@@ -1,16 +1,30 @@
 
 from models.kinematric_bicycle_model import Bicycle
 from controller import MpcController
-from path.path_generation import generate_random_path, generate_path
+from path.path_generation import generate_path
 from matplotlib import pyplot as plt
 import numpy as np
 from numpy import random
 import copy
 import pathlib
 import os
-from multiprocessing import Pool
+from dataclasses import dataclass
 import warnings
 warnings.filterwarnings("ignore", category=RuntimeWarning)
+
+@dataclass(init=True)
+class SimulationData:
+    seed: int
+    path : np.ndarray
+    path_contoller : np.ndarray
+    reference : np.ndarray
+    inputs : np.ndarray
+    measured_states : np.ndarray
+    tlf : float = 0.
+    tlr : float = 0.
+    lf : float = 0.
+    lr : float = 0.
+    input_ratio : float = 0.
 
 
 class Simulation:

### 2022-01-23 src/bo_mpc/visualize_simulation.py

@@ -1,0 +1,104 @@
+from simulation import SimulationData
+from visualize_gp import visualize_2d
+import json
+from matplotlib import pyplot as plt
+from matplotlib.widgets import Button
+import numpy as np
+import argparse
+import glob
+import os
+import re
+from path.path_generation import generate_path
+from models.kinematric_bicycle_model import Bicycle
+from numpy import random
+
+def load_from_file(file) -> SimulationData:
+    seed_re = "sim(\\d+).npz"
+    seed = int(re.search(seed_re, file).group(1))
+    if "l_estimation" in file:
+        l_re = re.search("l_estimation_(\\d+\\.\\d+)_(\\d+\\.\\d+)\\\\(\\d+\\.\\d+)_(\\d+\\.\\d+)", file)
+        tlr = float(l_re.group(1))
+        tlf = float(l_re.group(2))
+        lr = float(l_re.group(4))
+        lf = float(l_re.group(3))
+        ratio = 0.002
+
+    elif "input_ratio" in file:
+        l_re = re.search("input_ratio_(\\d+\\.\\d+)_(\\d+\\.\\d+)\\\\(\\d+\\.\\d+)\\\\(\\d+\\.\\d+)_(\\d+\\.\\d+)", file)
+        tlr = float(l_re.group(1))
+        tlf = float(l_re.group(2))
+        ratio = float(l_re.group(3))
+        lr = tlr
+        lf = tlf
+
+    npf = np.load(file)
+    if "ref" not in npf:
+        bike = Bicycle(tlr,tlf,dt=0.1)
+        random.seed(seed)
+        ref = generate_path(bike, 200)
+    else:
+        ref = npf["ref"]
+    sim = SimulationData(seed, npf["pos"], npf["pos_c"], ref, npf["u"], measured_states=[0, 1], tlf=tlf, tlr=tlr, lf=lf, lr=lr, input_ratio=ratio)
+    return sim
+
+if __name__ == "__main__":
+    parser = argparse.ArgumentParser(description="Script to visualize past simulations.")
+    parser.add_argument("glob_path")
+    args = parser.parse_args()
+
+    print(args.glob_path)
+    files = glob.glob(args.glob_path)#, root_dir=os.curdir)
+    print(files)
+
+    next_i = 0
+    class Index:
+        ind = 0
+        def __init__(self, files):
+            self.ind = 0
+            print(files[self.ind])
+            sim = load_from_file(files[self.ind])
+
[…]

### 2022-01-24 paper.tex
wip

@@
 \fancyfoot{} % Blank out the default footer
-\fancyhead[C]{Running title $\bullet$ May 2016 $\bullet$ Vol. XXI, No. 1} % Custom header text
+\fancyhead[C]{Bayesian Optimization in Model Predictive Control $\bullet$ Frederic Dlugi $\bullet$ January 2022} % Custom header text
 \fancyfoot[RO,LE]{\thepage} % Custom footer text
@@
 \posttitle{\end{center}} % Article title closing formatting
-\title{Baysian Optimization in MPC} % Article title
+\title{Bayesian Optimization in Model Predictive Control} % Article title
 \author{%
@@
 \begin{abstract}
-\noindent When using MPC to control a vehicle it is often neccesary to fine tune vehicle parameter for the model to match reality.
+\noindent When using MPC to control a vehicle it is often neccesary to fine tune control and vehicle parameters to get good performance in reality.
 This is a time consuming process that often relies on trial and error or grid search of the parameter space.
@@
 We simulate vehicle dynamics of a simple bicycle model with two parameters.
-We try to find the real vehicle parameters by optimizing controller performance with different parameters using Bayesian Optimization.
+We try to find the MPC cost ratios and real vehicle parameters by optimizing controller performance with different parameters using Bayesian Optimization.
 \end{abstract}
@@
 It is best suited for optimizing functions, where each evaluation takes a long time.
-The number of input dimensions for bayesian optimization is typically less then $20$. \cite{frazier2018tutorial}
+The number of input dimensions for bayesian optimization is typically less then $20$ \cite{frazier2018tutorial}.
 Bayesian Optimization uses an aquisition function that operates on a gaussian process of the evaluations.
@@
 
-We use a kinematic bicycle model defined by algorithm \ref{alg:kbm} and \ref{alg:ubm}. We try to estimate the $L_r$ and $L_f$ in this paper.
+We use a kinematic bicycle model defined by algorithm \ref{alg:kbm} and \ref{alg:ubm} \cite{kinematicBicycleModel}. We try to estimate the $L_r$ and $L_f$ in this paper.
 \begin{algorithm}
@@
     \begin{algorithmic}
-        \State $v \gets \text{speed (input)}$
+        \State $a \gets \text{accelaration (input)}$
         \State $\omega \gets \text{steering angle rate (input)}$
         \State $x_c, y_c \gets \text{center position}$
+        \State $v \gets \text{speed}$
         \State $\theta \gets \text{absolute angle}$
@@
     \begin{algorithmic}
+        \State $\dot{v} \gets a$
         \State $\dot{x_c} \gets v * \cos{(\theta + \beta)}$
@@
 
+\section{State of the art}
+In Multi-Objective Optimization of a Path-following MPC for Vehicle Guidance \cite{gharib2021multi} BO was used to weight error terms to find the ratio results in the best reference tracking performance of a complex vehicle model in a simulation.
+Using BO with expected improvement as acquisition function.
+They found the BO algorithm used to be more efficient then random search when comparing the number of evaluations needed to find satisfactory results.
+
+Lukas Hewing and colaborators \cite{hewing2018cautious} explored the use of BO to decrease model missmatch.
+They learned a dynamics model of a miniature race car using a gaussian process from real data, while the vehicle is controlled through MPC.
+They experimented with sparse approximations to gaussian processes to use them in real-time applications.
+
+What papers did similar things?
+What where their problems?
+What did they achieve?
+What do we do differently?
+
 \section{Experiments}
 
-Experiment Setup
+In this paper we explore two uses of BO in the MPC context.
+First we try to find the optimal ratio between input (accelaration and steering angle rate) cost and reference tracking ($x - x_{ref}$ and $y - y_{ref}$).
+Secondly we try to find vehicle specific parameters using BO.
+\subsection{Experiment setup}
+Show specific MPC implementation. Optimizer used is scipys SLSQP.
+Show cost function
+Show setup with graph
+Show generate function
+Name number of evaluations per BO step and number of BO steps.
+
+\subsection{Exploring MPC cost ratio using BO}
+Compare different acquisition functions.
+Cost ratio was explored with different vehicle parameters (show different gaussian processes)
+
+\subsection{Finding Vehicle parameters using BO}
+Using different target parameters
+Using different acquisition functions
+Show main target (1.2, 0.8) -> accumulate all evaluations in one gaussian process
+
+\section{Conclusion}
+There is no general ratio between input and reference cost. It depends on vehicle parameters. The optimal ratio is still similar.
+Finetuning still required. Using many different reference tracks, leads to better results.
+
+\section{Further research}
+Use more suffisticated MPC.
+Try more hyper parameters.
+Try if this works on real hardware or using one simple and one complex vehicle model.
+Try measuring more or less vehicle states.
+Add uncertency to measurents and model.
 

### 2022-02-07 paper.tex
wip

@@
 
-\section{State of the art}
-In Multi-Objective Optimization of a Path-following MPC for Vehicle Guidance \cite{gharib2021multi} BO was used to weight error terms to find the ratio results in the best reference tracking performance of a complex vehicle model in a simulation.
+\section{Literatur review}
+In Multi-Objective Optimization of a Path-following MPC for Vehicle Guidance \cite{gharib2021multi} BO is used to weight error terms to find the ratio results in the best reference tracking performance of a complex vehicle model in a simulation.
 Using BO with expected improvement as acquisition function.
@@
 
+Lu Qiugang and colaborators \cite{lu2020mpc} explored MPC parameter tuning in HVAC plants.
+They had good results and with limited evaluations.
+They used upper confidence bound as acquisition function and also had two tuning parameters.
+The experiment setup is similar to the one used in this paper.
+
+% Do i need more here?
 What papers did similar things?
@@
 Secondly we try to find vehicle specific parameters using BO.
+
 \subsection{Experiment setup}
+Show block diagram for both experiments. (Show setup with graph)
 Show specific MPC implementation. Optimizer used is scipys SLSQP.
 Show cost function
-Show setup with graph
 Show generate function
 Name number of evaluations per BO step and number of BO steps.
+
+To generate the path, we first generate random normal distributed inputs. Then we use the Bicycle model with real $L_r$ and $L_f$ values with these inputs to generate $N$ sequential positions.
+\begin{algorithm}
+    \caption{Generate Path}
+    \label{alg:path_gen}
+    \begin{algorithmic}
+        \State $L_r \gets \text{Distance center to rear wheel}$
+        \State $L_f \gets \text{Distance center to font wheel}$
+        \State $\text{initPos} \gets (0,0)$
+        \State $\text{N} \gets \text{Number of path points}$
+        \State $\text{bike} \gets \text{Bike(}L_r, L_f, \text{initPos}\text{)}$
+        \State $\text{inputs} \gets normalDistibutedArray(N - 1, 2)$
+        \State $\text{path} \gets zeroArray(N, 2)$
+        \For{$i= 1 \to (N-1)$}
+            \State $\text{bike}.step(\text{inputs[i - 1]})$
+            \State $\text{path[i]} \gets \text{bike}.getPosition()$
+        \EndFor
+        \State $return \text{ path}$
+    \end{algorithmic}
+\end{algorithm}
 
@@
 
-\section{Conclusion}
+\section{Conclusion and further research}
 There is no general ratio between input and reference cost. It depends on vehicle parameters. The optimal ratio is still similar.
@@
 
-\section{Further research}
 Use more suffisticated MPC.

### 2022-02-13 paper.tex
wip

@@
         \State $\mu, \sigma \gets \text{gp.predict(x)}$
-        \State $z \gets (\mu - y_{max} - xi)/\sigma$
-        \State $result \gets (\mu - y_{max} - xi) * \text{cdf(z)} + \sigma * \text{pdf(z)}$
+        \State $z \gets (\mu - y_{max} - \xi)/\sigma$
+        \State $result \gets (\mu - y_{max} - \xi) * \text{cdf(z)} + \sigma * \text{pdf(z)}$
     \end{algorithmic}
@@
         \State $\mu, \sigma \gets \text{gp.predict(x)}$
-        \State $z \gets (\mu - y_{max} - xi)/\sigma$
+        \State $z \gets (\mu - y_{max} - \xi)/\sigma$
         \State $result \gets \text{cdf(z)}$
@@
 
-% Do i need more here?
-What papers did similar things?
-What where their problems?
-What did they achieve?
-What do we do differently?
-
 \section{Experiments}
@@
 \subsection{Experiment setup}
-Show block diagram for both experiments. (Show setup with graph)
-Show specific MPC implementation. Optimizer used is scipys SLSQP.
-Show cost function
-Show generate function
-Name number of evaluations per BO step and number of BO steps.
 
-To generate the path, we first generate random normal distributed inputs. Then we use the Bicycle model with real $L_r$ and $L_f$ values with these inputs to generate $N$ sequential positions.
+We explore two uses of BO for MPC.
+Firstly we try to find the optimal ratio of input cost to state cost ratio (See Fig.
+\ref{fig:ratio_experiment}).
+Secondly we estimate $L_r$ and $L_f$ parameters of our controller (See Fig.
+\ref{fig:l_value_experiment}).
+
+\begin{figure}[h]
+    \caption{Ratio experiment}
+    \centering
+    \includegraphics[width=0.45\textwidth]{fig_ratio_experiment.pdf}
+    \label{fig:ratio_experiment}
+\end{figure}
+
+\begin{figure}[h]
+    \caption{L-Value experiment}
+    \centering
+    \includegraphics[width=0.45\textwidth]{fig_l_value_experiment.pdf}
+    \label{fig:l_value_experiment}
+\end{figure}
+
+We used a standard MPC implementation that uses the cost function that is specified here (Alg. \ref{alg:cost_function}). We use scipys SLSQP optimizer to minimize the cost function.
+\begin{algorithm}
+    \caption{Cost function}
+    \label{alg:cost_function}
+    \begin{algorithmic}
+        \State $U \gets \text{Array of inputs } (N \times 2)$
+        \State $X \gets \text{Array of postions } (N \times 2)$
+        \State $X_{ref} \gets \text{Array of reference postions } (N \times 2)$
+
+        \For{$i = 0 \to (N-1)$}
+            \State $cost + (X[i] - X_{ref}[i])^T Q (X[i] - X_{ref}[i])$
+            \State $cost + U[i]^T R U[i]$
+        \EndFor
+        \State $return \text{ cost}$
+    \end{algorithmic}
+\end{algorithm}
+
+To generate the path, we first generate random normal distributed inputs.
+Then we use the Bicycle model with real $L_r$ and $L_f$ values with these inputs to generate $N$ sequential positions (See Alg.
+\ref{alg:path_gen}).
+We use $200$ steps for all paths and all experiments.
 \begin{algorithm}
@@
 \subsection{Exploring MPC cost ratio using BO}
-Compare different acquisition functions.
-Cost ratio was explored with different vehicle parameters (show different gaussian processes)
+
+In this experiment we aim to find the optimal ratio between the $Q$ and $R$ matrix in the cost function (Alg \ref{alg:cost_function}).
+We try this for different vehicles that are differentiated in their L-parameter.
+Since a full search of the ratios shows that Ratios between $0.001$ and $0.003$ result in better performance (See Figure \ref{fig:cost_ratio_full}).
+Therefore we explore only this range in all further experiments.
+We use $40$ tracks per step and do $30$ steps for both experiments.
+Increasing the regularization $\alpha$ parameter of the internal GP for BO to $0.01$ improved performance.
+
+\begin{figure}[h]
+    \caption{GP of full cost ratio search (Higher is better)}
+    \centering
+    \includegraphics[width=0.45\textwidth]{fig_cost_ratio_full.pdf}
+    \label{fig:cost_ratio_full}
+\end{figure}
+
+\begin{figure}[h]
+    \caption{GP of detailed cost ratio search (Higher is better)}
+    \centering
+    \includegraphics[width=0.45\textwidth]{fig_cost_ratio_detail.pdf}
+    \label{fig:cost_ratio_detail}
+\end{figure}
+
+We found that the optimum ratio for all tested vehicles all with a length of $2m$ is close to $0.002$ (See Fig \ref{fig:cost_ratio_detail}).
+This is highly depended on the MPC implementation and vehicle dynamics.
 
 \subsection{Finding Vehicle parameters using BO}
-Using different target parameters
-Using different acquisition functions
-Show main target (1.2, 0.8) -> accumulate all evaluations in one gaussian process
+
+We aim to find the vehicle specific parameters $L_f$ and $L_r$ using BO.
+We try multiple parameters and different aquisition functions.
+Since evaluating $40$ tracks still lead to noisy cost we need to choose a suitable exploration-exploitation parameter in our acquisition function.
+For Expected Improvement \ref{alg:ei} and Probability of Improvement \ref{alg:poi} a $\xi = 0.5$ worked the best.
+For Upper Confidence Bound a $\kappa=2.576$ worked the best.
+Increasing the regularization $\alpha$ parameter of the internal GP for BO to $0.1$ improved performance.
+
+\begin{figure}[h]
+    \caption{Mean of GP from L-parameter search (Target $L_f = 0.8$ $L_r = 1.2$)}
+    \centering
+    \includegraphics[width=0.45\textwidth]{fig_l_estimation_full.pdf}
+    \label{fig:l_param_full}
+\end{figure}
 
 \section{Conclusion and further research}
-There is no general ratio between input and reference cost. It depends on vehicle parameters. The optimal ratio is still similar.
-Finetuning still required. Using many different reference tracks, leads to better results.
+There is no general ratio between input and reference cost.
+It depends on vehicle dynamics vehicle parameters and the specific MPC implementation.
+The optimal ratio are still similar and close to $0.002$ with our experiment.
+Therefore finetuning is still required and using many different reference tracks, leads to better results.\\
 
-Use more suffisticated MPC.
-Try more hyper parameters.
-Try if this works on real hardware or using one simple and one complex vehicle model.
-Try measuring more or less vehicle states.
-Add uncertency to measurents and model.
+Finding the vehicle parameters works well using BO if you tweak the BO parameter for exploration and exploitation tradeoff and set the regularization parameter of the internal GP for the BO to a suitable value.\\
+
+In further reseach we would use a more suffisticated MPC since using the SLSQP optimizer from python does not work well for some states.
+It would also be interesting to try to find more vehicle parameters.
+Also we would like to try if the BO approach works on real hardware or using one simple and one complex vehicle model this could also be done in simulation by using different vehicle models for the controller and the simulation.
+In this paper we always measure the current vehicle position and update this inside the controller therefore measuring more or less vehicle states and the effects on the optimization performance would be interesting.
+To translate this approach to a real vehicle you would need to add uncertency to measurents and to the model.
 

### 2022-02-13 src/bo_mpc/visualize_gp.py

@@ -3,11 +3,11 @@ from matplotlib import pyplot as plt
 from sklearn.gaussian_process import GaussianProcessRegressor
 from sklearn.gaussian_process.kernels import Matern
 
-def visualize_1d(X, y, bounds):
+def visualize_1d(X, y, bounds, color='b', show_observations=True, show_legend=True):
 
     gp = GaussianProcessRegressor(
             kernel=Matern(nu=2.5),
-            alpha=1e-6,
+            alpha= 0.1,
             normalize_y=True,
             n_restarts_optimizer=5,
             random_state=1,
@@ -19,19 +19,23 @@ def visualize_1d(X, y, bounds):
 
     mean, std = gp.predict(x_plot, return_std=True)
 
-    plt.plot(x_plot, mean, label="Mean")
-    plt.fill_between(x_plot.ravel(), mean - 1.96 * std, mean + 1.96 * std, alpha=0.5, label="95% confidence interval")
-    plt.scatter(X, y, label="Observations")
-    i_max = np.argmax(y)
-    plt.scatter(X[i_max], y[i_max], label="Found optimum")
-    plt.legend()
+    plt.plot(x_plot, mean, label="Mean", color=color)
+    plt.fill_between(x_plot.ravel(), mean - std, mean + std, alpha=0.2, label="$1 \sigma$ confidence interval", color=color)
+
+    if show_observations:
+        plt.scatter(X, y, marker="x",label="Observations", color=color)
+        i_max = np.argmax(y)
+        plt.scatter(X[i_max], y[i_max], marker="^", label="Optimum", color=color)
+
+    if show_legend:
+        plt.legend()
 
 
 def visualize_2d(X, y, bounds):
 
     gp = GaussianProcessRegressor(
             kernel=Matern(nu=2.5),
-            alpha=1e-6,
+            alpha=1e-1,
             normalize_y=True,
             n_restarts_optimizer=5,
             random_state=1,
@@ -49,7 +53,7 @@ def visualize_2d(X, y, bounds):
     img_mean = mean.reshape((500, 500))
 
     plt.contourf(img_mean, levels=20, extent=[*bounds[0], *bounds[1]])
-    plt.scatter(X[:,0], X[:,1], label="Observations")
+    plt.scatter(X[:,0], X[:,1], label="Observations", marker='x', color='r')
     i_max = np.argmax(y)
     plt.scatter(X[i_max, 0], X[i_max, 1], label="Found optimum")
     plt.legend()

### 2022-02-13 src/bo_mpc/visualize_input_ratio.py

@@ -2,23 +2,41 @@ from visualize_gp import visualize_1d
 import json
 from matplotlib import pyplot as plt
 import numpy as np
+import argparse
+import os
 
-if __name__ == "__main__":
-    json_file = open("logs.json")
+def directory(path):
+    if not os.path.isdir(path) and os.path.isfile(os.path.join(path, "logs.json")):
+        raise argparse.ArgumentTypeError(f"{path} is not a valid directory")
+    return path
 
-    x = []
-    y = []
-    for line in json_file.readlines():
-        json_l = json.loads(line)
-        x += [json_l["params"]["r"]]
-        y += [json_l["target"]]
+def main():
+    parser = argparse.ArgumentParser(prog="Visualize Cost Ratio Experiments")
+    parser.add_argument("paths", type=directory, nargs='+')
+    args = parser.parse_args()
 
-    x = np.array(x)
-    x = np.reshape(x, (-1, 1))
-    y = np.array(y)
+    colors = ["tab:blue", "tab:orange", "tab:green", "tab:red"]
 
-    print(x.shape, y.shape)
-    visualize_1d(x, y, (x.min(), x.max()))
+    plt.figure(figsize=(8,5))
+
+    for i, path in enumerate(args.paths):
+        json_file = open(os.path.join(path, "logs.json"))
+
+        x = []
+        y = []
+        for line in json_file.readlines():
+            json_l = json.loads(line)
+            x += [json_l["params"]["r"]]
+            y += [json_l["target"]]
+
+        x = np.array(x)
+        x = np.reshape(x, (-1, 1))
+        y = np.array(y)
+
+        visualize_1d(x, y, (x.min(), x.max()), colors[i], len(args.paths) == 1, len(args.paths) == 1)
     plt.xlabel("Ratio (input/output cost)")
     plt.ylabel("Accumulated cost")
     plt.show()
+
+if __name__ == "__main__":
+    main()

### 2022-02-13 src/bo_mpc/visualize_l_estimation.py

@@ -2,9 +2,20 @@ from visualize_gp import visualize_2d
 import json
 from matplotlib import pyplot as plt
 import numpy as np
+import argparse
+import os
 
-if __name__ == "__main__":
-    json_file = open("logs.json")
+def directory(path):
+    if not os.path.isdir(path) and os.path.isfile(os.path.join(path, "logs.json")):
+        raise argparse.ArgumentTypeError(f"{path} is not a valid directory")
+    return path
+
+def main():
+    parser = argparse.ArgumentParser(prog="Visualize Cost Ratio Experiments")
+    parser.add_argument("path", type=directory)
+    args = parser.parse_args()
+
+    json_file = open(os.path.join(args.path, "logs.json"))
 
     x = []
     y = []
@@ -20,3 +31,6 @@ if __name__ == "__main__":
     plt.xlabel("$L_f$")
     plt.ylabel("$L_r$")
     plt.show()
+
+if __name__ == "__main__":
+    main()

### 2022-03-19 paper.tex
wip

@@
 \usepackage{amsmath}
+
+\usepackage{todonotes}
 
@@
 \begin{abstract}
-\noindent When using MPC to control a vehicle it is often neccesary to fine tune control and vehicle parameters to get good performance in reality.
+\noindent When using MPC to control a vehicle it is often necessary to fine tune control and vehicle parameters to get good performance in reality.
 This is a time consuming process that often relies on trial and error or grid search of the parameter space.
@@
 \subsection{What is Bayesian Optimization?}
-Bayesian optimization is a strategy for global optimization of backbox functions. (Alg. \ref{alg:bo})
+Bayesian optimization is a strategy for global optimization of black-box functions. (Alg. \ref{alg:bo})
 It therefore does not require the computation of gradients, but works best on continuous functions.
@@
 The number of input dimensions for bayesian optimization is typically less then $20$ \cite{frazier2018tutorial}.
-Bayesian Optimization uses an aquisition function that operates on a gaussian process of the evaluations.
+Bayesian Optimization uses an acquisition function that operates on a gaussian process of the evaluations.
 In this paper we evaluate three different acquisition functions: Expected Improvement (Alg.
@@
 \subsection{What is MPC?}
-MPC is an acronym for model predicive control.
+MPC is an acronym for model predictive control.
 It replaces classical control algorithms that usually work in an offline manner with an optimizer, that solves the optimal control problem for a receding horizon (Figure \ref{fig:mpc}).
@@
 \begin{figure}[h]
-    \caption{Blockdiagram of MPC algorithm}
+    \caption{Block-diagram of MPC algorithm}
     \centering
@@
     \label{alg:ubm}
-    \begin{algorithmic}
+    \begin{algorithmic}[1]
         \State $\dot{v} \gets a$
@@
 
-\section{Literatur review}
+\section{Literature review}
 In Multi-Objective Optimization of a Path-following MPC for Vehicle Guidance \cite{gharib2021multi} BO is used to weight error terms to find the ratio results in the best reference tracking performance of a complex vehicle model in a simulation.
@@
 
-Lukas Hewing and colaborators \cite{hewing2018cautious} explored the use of BO to decrease model missmatch.
+Lukas Hewing and collaborators \cite{hewing2018cautious} explored the use of BO to decrease model mismatch.
 They learned a dynamics model of a miniature race car using a gaussian process from real data, while the vehicle is controlled through MPC.
@@
 
-Lu Qiugang and colaborators \cite{lu2020mpc} explored MPC parameter tuning in HVAC plants.
+Lu Qiugang and collaborators \cite{lu2020mpc} explored MPC parameter tuning in HVAC plants.
 They had good results and with limited evaluations.
@@
 In this paper we explore two uses of BO in the MPC context.
-First we try to find the optimal ratio between input (accelaration and steering angle rate) cost and reference tracking ($x - x_{ref}$ and $y - y_{ref}$).
+First we try to find the optimal ratio between input (acceleration and steering angle rate) cost and reference tracking ($x - x_{ref}$ and $y - y_{ref}$).
 Secondly we try to find vehicle specific parameters using BO.
@@
 
-We used a standard MPC implementation that uses the cost function that is specified here (Alg. \ref{alg:cost_function}). We use scipys SLSQP optimizer to minimize the cost function.
+We used a standard MPC implementation that uses the cost function that is specified here (Alg. \ref{alg:cost_function}). We use scipy's SLSQP optimizer to minimize the cost function.
 \begin{algorithm}
@@
 We aim to find the vehicle specific parameters $L_f$ and $L_r$ using BO.
-We try multiple parameters and different aquisition functions.
+We try multiple parameters and different acquisition functions.
 Since evaluating $40$ tracks still lead to noisy cost we need to choose a suitable exploration-exploitation parameter in our acquisition function.
@@
 The optimal ratio are still similar and close to $0.002$ with our experiment.
-Therefore finetuning is still required and using many different reference tracks, leads to better results.\\
+Therefore fine tuning is still required and using many different reference tracks, leads to better results.
 
-Finding the vehicle parameters works well using BO if you tweak the BO parameter for exploration and exploitation tradeoff and set the regularization parameter of the internal GP for the BO to a suitable value.\\
+Finding the vehicle parameters works well using BO if you tweak the BO parameter for exploration and exploitation tradeoff and set the regularization parameter of the internal GP for the BO to a suitable value.
+Since we only estimated two parameters in our experiment using BO in this case was not helpful.
 
-In further reseach we would use a more suffisticated MPC since using the SLSQP optimizer from python does not work well for some states.
+In further research we would use a more sophisticated MPC since using the SLSQP optimizer from python does not work well for some states.
 It would also be interesting to try to find more vehicle parameters.
@@
 In this paper we always measure the current vehicle position and update this inside the controller therefore measuring more or less vehicle states and the effects on the optimization performance would be interesting.
-To translate this approach to a real vehicle you would need to add uncertency to measurents and to the model.
+To translate this approach to a real vehicle you would need to add uncertainly to measurements and to the model.
 

### 2022-03-19 src/bo_mpc/visualize_simulation.py

@@ -16,7 +16,7 @@ def load_from_file(file) -> SimulationData:
     seed_re = "sim(\\d+).npz"
     seed = int(re.search(seed_re, file).group(1))
     if "l_estimation" in file:
-        l_re = re.search("l_estimation_(\\d+\\.\\d+)_(\\d+\\.\\d+)\\\\(\\d+\\.\\d+)_(\\d+\\.\\d+)", file)
+        l_re = re.search("l_estimation_(\\d+\\.\\d+)_(\\d+\\.\\d+).*\\\\(\\d+\\.\\d+)_(\\d+\\.\\d+)", file)
         tlr = float(l_re.group(1))
         tlf = float(l_re.group(2))
         lr = float(l_re.group(4))

### 2022-03-19 paper.tex
update intro

@@
 \begin{abstract}
-\noindent When using MPC to control a vehicle it is often necessary to fine tune control and vehicle parameters to get good performance in reality.
+\noindent When using MPC to control a vehicle it is often necessary to fine tune model parameters and control costs to get good performance in reality.
 This is a time consuming process that often relies on trial and error or grid search of the parameter space.
 In this paper we evaluate the use of Bayesian Optimization to tackle this problem.
-We simulate vehicle dynamics of a simple bicycle model with two parameters.
-We try to find the MPC cost ratios and real vehicle parameters by optimizing controller performance with different parameters using Bayesian Optimization.
+We simulate vehicle dynamics of a simple bicycle model with two model parameters.
+We try to find the MPC cost matrices and real vehicle parameters by optimizing controller performance with different parameters using Bayesian Optimization.
 \end{abstract}
@@
 
-\lettrine[nindent=0em,lines=3]{W} e start by introducing bayesian optimization and MPC on a high level.
-This is done to let you know what I understand when talking about these algorithms.
+\lettrine[nindent=0em,lines=3]{W}  hen using Model Predictive Control (MPC) to solve a problem a lot of fine tuning and calibration needs to be done for an algorithm to work on a plant.
+Assuming the popular quadratic cost function the cost matrices for input and state of the model need to be chosen.
+Many models also have parameters that represent physical aspects of the plant for best performance the model parameters need to match the plant.
+In many cases the model is a simplification of the plant, therefore multiple physical features may influence the optimal parameters of the model and can't be easily measured.
+In the design process of a MPC-controller the selection of optimal parameters and cost matrices needs deep understanding of the plant and time for manual tuning of the variables.
+In this paper we explore to use of Bayesian Optimization (BO) for the optimal selection of cost matrices and model parameters using a simulated bicycle vehicle.
 
@@
 \subsection{What is MPC?}
-MPC is an acronym for model predictive control.
-It replaces classical control algorithms that usually work in an offline manner with an optimizer, that solves the optimal control problem for a receding horizon (Figure \ref{fig:mpc}).
+MPC replaces classical control algorithms that usually work in an offline manner with an optimizer, that solves the optimal control problem for a receding horizon (Figure \ref{fig:mpc}).
 Each action taken is the first action of the optimal control plan.
